---
title: "benchmark mRNA mixtures"
output: html_notebook
---

# Preparation

Load some packages

```{r message=FALSE}
library(scran)
library(scater)
library(ggplot2)
library(edgeR)
library(DESeq)
# library(ggridges)
```

Load R data

```{r}
# load("SCE_sc.RData")
load("SCE.RData")
```

2: CEL-Seq2
8: SORT-Seq


Assign group information

```{r}
sce2_qc$group <- paste(sce2_qc$H1975_prop, sce2_qc$H2228_prop, sce2_qc$HCC827_prop, sep = "-")
sce8_qc$group <- paste(sce8_qc$H1975_prop, sce8_qc$H2228_prop, sce8_qc$HCC827_prop, sep = "-")
```

Function for calculate silhouette width of PCA analysis, group samples according to their cell type

```{r}
library(cluster)

calcSilhouette <- function(SCE, name){
  sil <- silhouette(as.numeric(factor(SCE$group)), dist(reducedDim(SCE)))
  sil <- data.frame(sil[1:nrow(sil),])
  method <- rep(name, nrow(sil))
  sil <- cbind(sil, method)
}
```

<!-- ```{r} -->
<!-- cor_mRNA_within <- function(sce, norm_assay = "norm_exprs") { -->
<!--   cor_val = c() -->
<!--   cell = c() -->
<!--   norm_mat = assay(sce, norm_assay) -->
<!--   for(group in unique(sce$group)){ -->
<!--     sel_cell = colData(sce)$group == group -->
<!--     if(table(sel_cell)["TRUE"] > 1){ -->
<!--     cor_mat = cor(norm_mat[,sel_cell]) -->
<!--     cor_mat[!lower.tri(cor_mat)] = NA -->
<!--     cor_mat = as.numeric(cor_mat) -->
<!--     cor_mat = cor_mat[!is.na(cor_mat)] -->
<!--     cor_val = c(cor_val, cor_mat) -->
<!--     cell = c(cell, rep(group, length(cor_mat))) -->
<!--     } -->
<!--   } -->
<!--   res_df = data.frame(cor_val = cor_val, group = cell) -->
<!-- } -->
<!-- ``` -->



```{r}
# require the SCE to be the result of dimension reduction
# this function calculates the true distance of the k cells that is closest in the dimension reduction result where k is the number of cell within  group
cal_knn_avg_distance <- function(SCE){
  cell_annotation = data.frame(colData(SCE))
  cell_matrix <- as.matrix(cell_annotation[,c("H1975_prop", "H2228_prop", "HCC827_prop")])
  true_distance_matrix = as.matrix(dist(cell_matrix, method = "euclidean"))
  rd_distance_matrix = as.matrix(dist(reducedDim(SCE)))
  
  avg_distance = c()
  grp = table(colData(SCE)$group)
  for(i in 1:ncol(SCE)){
    knn = order(rd_distance_matrix[rownames(cell_annotation)[i],])[1:(grp[cell_annotation$group[i]])]
    avg_distance = c(avg_distance, sum(true_distance_matrix[rownames(cell_annotation)[i], knn])/grp[cell_annotation$group[i]])
  }
  names(avg_distance) = rownames(cell_annotation)
  return(avg_distance)
}
  
```

# Benchmarking for normalization methods

data set: CEL-seq2 data

## raw log count

```{r}
time_raw <- system.time(
logcounts(sce2_qc) <- log2(counts(sce2_qc) + 1))
```

## log cpm

```{r}
time_CPM <- system.time(
{assay(sce2_qc, "norm_cpm") = log2(cpm(counts(sce2_qc)) + 1)})
```

## scran

```{r}
time_scran <- system.time({
  sce2_qc <- computeSumFactors(sce2_qc)
  sce2_qc <- normalise(sce2_qc)
  assay(sce2_qc, "scran") = logcounts(sce2_qc)
})

```

## TMM from edgeR

```{r}
time_TMM <- system.time({
  sizeFactors(sce2_qc) <- calcNormFactors(counts(sce2_qc), method = "TMM")
  sce2_qc <- normalise(sce2_qc)
  assay(sce2_qc, "TMM") <- logcounts(sce2_qc)
})
```

## DESeq

```{r}
time_DESeq <- system.time({
  sizeFactors(sce2_qc) <- estimateSizeFactorsForMatrix(counts(sce2_qc))
  sce2_qc <- normalize(sce2_qc)
  assay(sce2_qc, "DESeq") <- logcounts(sce2_qc)
})
detach("package:DESeq", unload=TRUE) 
```

## SCnorm

```{r}
library(SCnorm)

time_SCnorm <- system.time({
  SCnorm_out <- SCnorm(Data = counts(sce2_qc), Conditions = as.numeric(factor(sce2_qc$group)), PrintProgressPlots = T)
  assay(sce2_qc, "SCnorm") = log2(metadata(SCnorm_out)$NormalizedData + 1)
})
save(SCnorm_out, file = "SCnorm_out_mRNAmix.RData")
detach("package:SCnorm", unload=TRUE) 
```

## Linnorm

```{r}
library(Linnorm)
time_Linnorm <- system.time({
linnorm_normalized <- Linnorm(counts(sce2_qc))
assay(sce2_qc, "linnorm") <- linnorm_normalized
})
detach("package:Linnorm", unload=TRUE) 

```

## scone

```{r}
library(scone)

expr <- counts(sce2_qc)
bio <- factor(sce2_qc$group)


#Read in housekeeping gene list and convert gene symbol to ensembl gene ID using biomaRt.


hkgene <- read.csv("h-scHKgenes.csv", stringsAsFactors = FALSE, col.names = 1)
library(biomaRt)
ensembl = useMart("ensembl")
ensembl = useDataset("hsapiens_gene_ensembl", mart = ensembl)
hk <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"), filters = "hgnc_symbol", mart = ensembl, values = hkgene$X1)


#Use housekeeping genes as negetive control
time_scone <- system.time({
  negcon <- intersect(rownames(expr), hk$ensembl_gene_id)
  # creating a SconeExperiment Object
  my_scone <- SconeExperiment(expr, bio = bio, negcon_ruv = rownames(expr) %in% negcon)
  
  scaling=list(none=identity, # Identity - do nothing
               
               # eff = EFF_FN, # User-defined function
               
               sum = SUM_FN, # SCONE library wrappers...
               tmm = TMM_FN, 
               uq = UQ_FN,
               fq = FQT_FN,
               deseq = DESEQ_FN)
  
  BiocParallel::register(
    BiocParallel::SerialParam()
  ) # Register BiocParallel Serial Execution
  
  my_scone <- scone(my_scone,
                    #imputation = imputation, impute_args = impute_args,
                    scaling=scaling,
                    run=TRUE,
                    eval_kclust = 2:6,stratified_pam = TRUE,
                    k_qc=0, k_ruv = 3,
                    adjust_bio="no",
                    return_norm = "in_memory",
                    zero = "postadjust")
  # NO RUV
  
  

  
  scores <- get_scores(my_scone)
  score_ranks <- get_score_ranks(my_scone)
  out_norm = get_normalized(my_scone,method = rownames(get_params(my_scone))[1])
  
  assay(sce2_qc, "scone") = log2(out_norm + 1)
})

detach("package:scone", unload=TRUE) 
detach("package:biomaRt", unload=TRUE) 
```

## BASiCS

```{r}
load("benchmark_mRNA_BASiCS.RData")
DenoisedCounts <- BASiCS_DenoisedCounts(Data = sce2_qc_B, Chain = chain)
DenoisedCounts = DenoisedCounts[rownames(sce2_qc_B), colnames(sce2_qc_B)]
assay(sce2_qc_B, "BASiCS") = log2(DenoisedCounts + 1)
assay(sce2_qc, "BASiCS") = assay(sce2_qc_B, "BASiCS")[match(rownames(sce2_qc), rownames(sce2_qc_B)),] 
rm(sce2_qc_B)
```

## zinbwave

```{r}
load("zinbwave_mRNA.RData")
assay(sce2_qc, "zinbwave") <- assay(se_norm, "normalizedValues")

```



## knn

test the influence of different k value
```{r}
source("/wehisan/home/allstaff/d/dong.x/mixture/knn-smoothing-master/knn_smooth.R")
#time1 <- Sys.time()
time_knn1 <- system.time({
sce2_smoothed = knn_smoother(counts(sce2_qc), k = 1)
assay(sce2_qc, "knn1") = log2(sce2_smoothed + 1)
})
# time2 <- Sys.time()
# time_knn1 <- c("knn1", time2 - time1)

# time1 <- Sys.time()
time_knn3 <- system.time({
sce2_smoothed = knn_smoother(counts(sce2_qc), k = 3)
assay(sce2_qc, "knn3") = log2(sce2_smoothed + 1)
})
# time2 <- Sys.time()
# time_knn3 <- c("knn3", time2 - time1)

# time1 <- Sys.time()
time_knn7 <- system.time({
sce2_smoothed = knn_smoother(counts(sce2_qc), k = 7)
assay(sce2_qc, "knn7") = log2(sce2_smoothed + 1)
})
# time2 <- Sys.time()
# time_knn7 <- c("knn7", time2 - time1)

# time1 <- Sys.time()
time_knn15 <- system.time({
sce2_smoothed = knn_smoother(counts(sce2_qc), k = 15)
assay(sce2_qc, "knn15") = log2(sce2_smoothed + 1)
})
# time2 <- Sys.time()
# time_knn15 <- c("knn15", time2 - time1)

# time1 <- Sys.time()
time_knn31 <- system.time({
sce2_smoothed = knn_smoother(counts(sce2_qc), k = 31)
assay(sce2_qc, "knn31") = log2(sce2_smoothed + 1)
})
# time2 <- Sys.time()
# time_knn31 <- c("knn31", time2 - time1)

# time1 <- Sys.time()
time_knn63 <- system.time({
sce2_smoothed = knn_smoother(counts(sce2_qc), k = 63)
assay(sce2_qc, "knn63") = log2(sce2_smoothed + 1)
})
# time2 <- Sys.time()
# time_knn63 <- c("knn63", time2 - time1)

# time1 <- Sys.time()
time_knn127 <- system.time({
sce2_smoothed = knn_smoother(counts(sce2_qc), k = 127)
assay(sce2_qc, "knn127") = log2(sce2_smoothed + 1)
})
# time2 <- Sys.time()
# time_knn127 <- c("knn127", time2 - time1)

# time1 <- Sys.time()
time_knn255 <- system.time({
sce2_smoothed = knn_smoother(counts(sce2_qc), k = 255)
assay(sce2_qc, "knn255") = log2(sce2_smoothed + 1)
})
# time2 <- Sys.time()
# time_knn255 <- c("knn255", time2 - time1)
```

## DrImpute

```{r}
library(DrImpute)
time_DrImpute <- system.time({
  imp <- DrImpute(assay(sce2_qc, "scran"))
  assay(sce2_qc, "scran_DrImpute") <- imp
})
```

## SAVER

```{r}
library(SAVER)
time_SAVER <- system.time({
registerDoParallel(cores = 64)
sce2_qc <- computeSumFactors(sce2_qc)
saver_out <- saver(counts(sce2_qc), size.factor = sizeFactors(sce2_qc))
#load("saver_out.RData")
assay(sce2_qc, "SAVER_scran") <- log2(saver_out$estimate + 1)
})
```
<!-- Timing stopped at: 3.089e+05 2.199e+04 6753 -->
<!-- 
Error in out[[i]][lasso.genes, ] <- matrix(unlist(tempvec), nrow = length(tempvec), : number of items to replace is not a multiple of replacement length
2.
saver(counts(sce2_qc), size.factor = sizeFactors(sce2_qc), parallel = TRUE)
1.
system.time({ registerDoParallel(cores = 64) sce2_qc <- computeSumFactors(sce2_qc) saver_out <- saver(counts(sce2_qc), size.factor = sizeFactors(sce2_qc), ...

-->


## PCA

```{r}
logcounts(sce2_qc) <- log2(counts(sce2_qc) + 1)
# sce2_qc_B$group <- paste(sce2_qc_B$H1975_prop, sce2_qc_B$H2228_prop, sce2_qc_B$HCC827_prop, sep = "-")

PCA1 <- runPCA(sce2_qc, exprs_values = "logcounts", ncomponents = 2)
PCA2 <- runPCA(sce2_qc, exprs_values = "norm_cpm", ncomponents = 2)
PCA3 <- runPCA(sce2_qc, exprs_values = "scran", ncomponents = 2) 
# PCA4 <- runPCA(sce2_qc, exprs_values = "SCnorm", ncomponents = 2) 
PCA5 <- runPCA(sce2_qc, exprs_values = "TMM", ncomponents = 2) 
PCA6 <- runPCA(sce2_qc, exprs_values = "DESeq", ncomponents = 2)
# PCA7 <- runPCA(sce2_qc, exprs_values = "knn15", ncomponents = 2)
PCA8 <- runPCA(sce2_qc, exprs_values = "knn31", ncomponents = 2) 
PCA9 <- runPCA(sce2_qc, exprs_values = "BASiCS", ncomponents = 2)
PCA10 <- runPCA(sce2_qc, exprs_values = "scone", ncomponents = 2)
PCA11 <- runPCA(sce2_qc, exprs_values = "scran_DrImpute", ncomponents = 2)
PCA12 <- runPCA(sce2_qc, exprs_values = "SAVER_scran", ncomponents = 2)
PCA13 <- runPCA(sce2_qc, exprs_values = "linnorm", ncomponents = 2)
PCA14 <- runPCA(sce2_qc, exprs_values = "zinbwave", ncomponents = 2)

col_mRNA <- rgb(sce2_qc$H1975_prop, sce2_qc$H2228_prop, sce2_qc$HCC827_prop)
# col_mRNA_B <- rgb(sce2_qc_B$H1975_prop, sce2_qc_B$H2228_prop, sce2_qc_B$HCC827_prop)

p1 <- plotPCASCE(PCA1, colour_by = "group",  exprs_values = "logcounts", ncomponents = 2) + ggtitle("log raw counts")+  scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))+ theme(legend.position="none") 
p2 <- plotPCASCE(PCA2, colour_by = "group",  exprs_values = "norm_cpm", ncomponents = 2) + ggtitle("CPM")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
p3 <- plotPCASCE(PCA3, colour_by = "group",  exprs_values = "scran", ncomponents = 2) + ggtitle("scran")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
# p4 <- plotPCASCE(PCA4, colour_by = "group", exprs_values = "SCnorm", ncomponents = 2) + ggtitle("SCnorm")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
p5 <- plotPCASCE(PCA5, colour_by = "group",  exprs_values = "TMM", ncomponents = 2) + ggtitle("TMM")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
p6 <- plotPCASCE(PCA6, colour_by = "group",  exprs_values = "DESeq", ncomponents = 2) + ggtitle("DESeq")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
# p7 <- plotPCASCE(PCA7, colour_by = "group", exprs_values = "knn15", ncomponents = 2) + ggtitle("kNN (k=15)")+ theme(legend.position="none")+ scale_fill_manual(values = c("#FF0000", "#00FF00", "#0000FF"), limits = c("H1975", "H2228", "HCC827"))
p8 <- plotPCASCE(PCA8, colour_by = "group",  exprs_values = "knn31", ncomponents = 2) + ggtitle("kNN (k=31)")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
p9 <- plotPCASCE(PCA9, colour_by = "group",  exprs_values = "BASiCS", ncomponents = 2) + ggtitle("BASiCS")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
p10 <- plotPCASCE(PCA10, colour_by = "group",  exprs_values = "scone", ncomponents = 2) + ggtitle("scone")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
p11 <- plotPCASCE(PCA11, colour_by = "group", exprs_values = "scran_DrImpute", ncomponents = 2) + ggtitle("scran+DrImpute")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
p12 <- plotPCASCE(PCA12, colour_by = "group", exprs_values = "SAVER_scran", ncomponents = 2) + ggtitle("scran+SAVER")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
p13 <- plotPCASCE(PCA13, colour_by = "group", exprs_values = "linnorm", ncomponents = 2) + ggtitle("Linnorm")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
p14 <- plotPCASCE(PCA14, colour_by = "group",  exprs_values = "zinbwave", ncomponents = 2) + ggtitle("ZINB-WaVE")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))


pdf("benchmark/thesis/PCA_mRNA_CEL-seq.pdf", height = 9)

multiplot( p1, p2, p3, p5, p6,  p8, p9, p10,   p13,p14, cols = 3)
dev.off()

pdf("benchmark/thesis/PCAz.pdf", width = 5, height = 5)
p14
dev.off()

pdf("benchmark/thesis/PCAl.pdf", width = 5, height = 5)
p13
dev.off()
```

## silhouette width

```{r}
sil1 <- calcSilhouette(PCA1, "raw count")
sil2 <- calcSilhouette(PCA2, "CPM")
sil3 <- calcSilhouette(PCA3, "scran")
# sil4 <- calcSilhouette(PCA4, "SCnorm")
sil5 <- calcSilhouette(PCA5, "TMM")
sil6 <- calcSilhouette(PCA6, "DESeq")
# sil7 <- calcSilhouette(PCA7, "kNN (k=15)")
sil8 <- calcSilhouette(PCA8, "kNN (k=31)")
sil9 <- calcSilhouette(PCA9, "BASiCS")
sil10 <- calcSilhouette(PCA10, "scone")
sil11 <- calcSilhouette(PCA11, "scran+DrImpute")
sil12 <- calcSilhouette(PCA12, "scran+SAVER")
sil13 <- calcSilhouette(PCA13, "Linnorm")
sil14 <- calcSilhouette(PCA14, "zinbwave")




se_norm$group <- sce2_qc$group
sil15 <- calcSilhouette(se_norm, "zinbwave W")

sil_all <- rbind(sil1, sil2, sil3, sil5, sil6, sil8, sil9, sil10, sil11, sil12, sil13, sil14, sil15)

# pdf("benchmark/silhouette_mRNA.pdf")
ggplot(sil_all, aes(x = reorder(method, sil_width), y=sil_width, fill=reorder(method, sil_width))) + geom_boxplot() + stat_summary(fun.y=mean, geom="point",size=2) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(legend.position = "none")+ labs(x="method", y="silhouette width")
# dev.off()
```


### timer

box plot coloured by time elapsed
```{r}
library(viridis)
time_all <- rbind(time_CPM, time_DESeq,  time_knn1, time_knn3, time_knn7, time_knn15, time_knn31, time_knn63, time_knn127, time_knn255, time_raw, time_scone, time_scran, time_TMM, time_BASiCS, time_Linnorm, time_zinbwave, time_DrImpute, time_SAVER)
time_all <- data.frame(time_all)
time_all$method <- c("CPM", "DESeq",  "kNN (k=1)", "kNN (k=3)", "kNN (k=7)", "kNN (k=15)", "kNN (k=31)", "kNN (k=63)", "kNN (k=127)", "kNN (k=255)", "raw count",  "scone", "scran", "TMM", "BASiCS", "Linnorm", "zinbwave", "scran+DrImpute", "scran+SAVER")
time_all_sil <- merge(sil_all, time_all, by = "method")

write.csv(time_all, "benchmark/mRNA_time_all.csv")
```



```{r}
meanSil <- sapply(unique(time_all_sil$method), function(x){
  mean(time_all_sil[time_all_sil$method == x, "sil_width"])
}, simplify = TRUE)
medSil <- sapply(unique(time_all_sil$method), function(x){
  median(time_all_sil[time_all_sil$method == x, "sil_width"])
}, simplify = TRUE)
meanSil <- data.frame(meanSil, medSil, method = unique(time_all_sil$method))
meanSil <- merge(meanSil, time_all, by = "method")

write.csv(meanSil, "benchmark/mRNA_meanSil.csv")
```

```{r}
# pdf("benchmark/medsiltime_mRNA.pdf", height=4)
ggplot(meanSil, aes(log10(elapsed), medSil, colour= reorder(method, medSil)))  + geom_text(aes(label=method)) + labs(x="log10 running time lapse", y="median silhouette width") +theme_bw()+ theme(legend.position = "none")
# dev.off()
```




## cor_df

```{r}
cor_mRNAmix_within = function(sce,norm_assay="norm_exprs"){
  cor_val = c()
  mRNA_amount = c()
  norm_mat = assay(sce, norm_assay)
  for(mix in 1:8){
    sel_cell = (colData(sce)$mix == mix) & (colData(sce)$mRNA_amount == 30)
    cor_mat = cor(norm_mat[,sel_cell])
    cor_mat[!lower.tri(cor_mat)] = NA
    cor_mat = as.numeric(cor_mat)
    cor_mat = cor_mat[!is.na(cor_mat)]
    
    cor_val = c(cor_val, cor_mat)
    mRNA_amount = c(mRNA_amount,rep(30,length(cor_mat)))
    
    sel_cell = (colData(sce)$mix == mix) & (colData(sce)$mRNA_amount == 15)
    cor_mat = cor(norm_mat[,sel_cell])
    cor_mat[!lower.tri(cor_mat)] = NA
    cor_mat = as.numeric(cor_mat)
    cor_mat = cor_mat[!is.na(cor_mat)]
    
    cor_val = c(cor_val, cor_mat)
    mRNA_amount = c(mRNA_amount,rep(15,length(cor_mat)))

    sel_cell = (colData(sce)$mix == mix) & (colData(sce)$mRNA_amount == 7.5)
    cor_mat = cor(norm_mat[,sel_cell])
    cor_mat[!lower.tri(cor_mat)] = NA
    cor_mat = as.numeric(cor_mat)
    cor_mat = cor_mat[!is.na(cor_mat)]
    
    cor_val = c(cor_val, cor_mat)
    mRNA_amount = c(mRNA_amount,rep(7.5,length(cor_mat)))
    
    sel_cell = (colData(sce)$mix == mix) & (colData(sce)$mRNA_amount == 3.75)
    cor_mat = cor(norm_mat[,sel_cell])
    cor_mat[!lower.tri(cor_mat)] = NA
    cor_mat = as.numeric(cor_mat)
    cor_mat = cor_mat[!is.na(cor_mat)]
    
    cor_val = c(cor_val, cor_mat)
    mRNA_amount = c(mRNA_amount,rep(3.75,length(cor_mat)))
  }
  res_df = data.frame(cor_val=cor_val,mRNA_amount=mRNA_amount)
}
```

```{r}
exprs <- c("logcounts", "norm_cpm","scran", "TMM", "DESeq", "knn31", "BASiCS",  "scone",  "scran_DrImpute", "linnorm",  "zinbwave", "SAVER_scran")
methods <- c("raw count", "CPM","scran", "TMM", "DESeq", "kNN", "BASiCS",  "scone",  "scran+DrImpute", "linnorm",  "ZINB-WaVE", "scran+SAVER")

cor_df <- sapply(exprs, function(x){ cor_mRNAmix_within(sce2_qc, norm_assay = x)})

cor_df_all <- data.frame()

for(i in 1:ncol(cor_df)){
  tmp <- data.frame(cor_df[,i])
  tmp$method <- methods[i]
  cor_df_all <- rbind(cor_df_all, tmp)
}

write.csv(cor_df_all, "benchmark/cor_df.csv")
```

```{r}
sil_all$method <- sub("zinbwave", "ZINB-WaVE", sil_all$method)
sil_all$method <- sub("kNN (k=31)", "kNN", sil_all$method)

```

```{r}
ggplot(cor_df_all, aes(reorder(method, cor_val), cor_val, fill = factor(mRNA_amount))) + geom_boxplot() + theme_bw() + scale_fill_brewer(palette = "Set2")
```

# save

```{r}
write.csv(sil_all, "benchmark/mRNA_sil_all.csv")

```

# Imputation
```{r}
library(DrImpute)
#Linnorm
imp <- DrImpute(assay(sce2_qc, "linnorm"))
assay(sce2_qc, "linnorm_DrImpute") <- imp
#scone
imp <- DrImpute(assay(sce2_qc, "scone"))
assay(sce2_qc, "scone_DrImpute") <- imp
#BASiCS
imp <- DrImpute(assay(sce2_qc, "BASiCS"))
assay(sce2_qc, "BASiCS_DrImpute") <- imp
#DESeq
imp <- DrImpute(assay(sce2_qc, "DESeq"))
assay(sce2_qc, "DESeq_DrImpute") <- imp
#CPM
imp <- DrImpute(assay(sce2_qc, "norm_cpm"))
assay(sce2_qc, "CPM_DrImpute") <- imp
#TMM
imp <- DrImpute(assay(sce2_qc, "TMM"))
assay(sce2_qc, "TMM_DrImpute") <- imp
#scran
imp <- DrImpute(assay(sce2_qc, "scran"))
assay(sce2_qc, "scran_DrImpute") <- imp
#zinbwave
imp <- DrImpute(assay(sce2_qc, "zinbwave"))
assay(sce2_qc, "zinbwave_DrImpute") <- imp 
```

```{r}

PCA_imp1 <- runPCA(sce2_qc, exprs_values = "linnorm_DrImpute")
PCA_imp2 <- runPCA(sce2_qc, exprs_values = "scone_DrImpute")
PCA_imp3 <- runPCA(sce2_qc, exprs_values = "BASiCS_DrImpute")
PCA_imp4 <- runPCA(sce2_qc, exprs_values = "DESeq_DrImpute")
PCA_imp5 <- runPCA(sce2_qc, exprs_values = "CPM_DrImpute")
PCA_imp6 <- runPCA(sce2_qc, exprs_values = "TMM_DrImpute")
PCA_imp7 <- runPCA(sce2_qc, exprs_values = "scran_DrImpute")
PCA_imp8 <- runPCA(sce2_qc, exprs_values = "zinbwave_DrImpute")

p_imp1 <- plotPCASCE(PCA_imp1, colour_by = "group", exprs_values = "linnorm_DrImpute", ncomponents = 2) + ggtitle("linnorm_DrImpute")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
p_imp2 <- plotPCASCE(PCA_imp2, colour_by = "group", exprs_values = "scone_DrImpute", ncomponents = 2) + ggtitle("scone_DrImpute")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
p_imp3 <- plotPCASCE(PCA_imp3, colour_by = "group", exprs_values = "BASiCS_DrImpute", ncomponents = 2) + ggtitle("BASiCS_DrImpute")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
p_imp4 <- plotPCASCE(PCA_imp4, colour_by = "group", exprs_values = "DESeq_DrImpute", ncomponents = 2) + ggtitle("DESeq_DrImpute")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
p_imp5 <- plotPCASCE(PCA_imp5, colour_by = "group", exprs_values = "CPM_DrImpute", ncomponents = 2) + ggtitle("CPM_DrImpute")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
p_imp6 <- plotPCASCE(PCA_imp6, colour_by = "group", exprs_values = "TMM_DrImpute", ncomponents = 2) + ggtitle("TMM_DrImpute")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
p_imp7 <- plotPCASCE(PCA_imp7, colour_by = "group", exprs_values = "scran_DrImpute", ncomponents = 2) + ggtitle("scran_DrImpute")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
p_imp8 <- plotPCASCE(PCA_imp8, colour_by = "group", exprs_values = "zinbwave_DrImpute", ncomponents = 2) + ggtitle("ZINB-WaVE_DrImpute")+ theme(legend.position="none")+ scale_fill_manual(values = unique(col_mRNA), limits = unique(sce2_qc$group))
# pdf("benchmark/mRNA_impu.pdf", width = 9)
multiplot( p_imp1, p_imp2, p_imp3, p_imp4, p_imp5, p_imp6, p_imp7,p_imp8, cols=3)
# dev.off()
```

```{r}
sil_imp_1 <- calcSilhouette(PCA_imp1, "linnorm+DrImpute")
sil_imp_2 <- calcSilhouette(PCA_imp2, "scone+DrImpute")
sil_imp_3 <- calcSilhouette(PCA_imp4, "BASiCS+DrImpute")
sil_imp_4 <- calcSilhouette(PCA_imp5, "DESeq+DrImpute")
sil_imp_5 <- calcSilhouette(PCA_imp6, "CPM+DrImpute")
sil_imp_6 <- calcSilhouette(PCA_imp7, "TMM+DrImpute")
sil_imp_7 <- calcSilhouette(PCA_imp6, "scran+DrImpute")
sil_imp_8 <- calcSilhouette(PCA_imp7, "ZINB-WaVE+DrImpute")

sil_imp <- rbind(sil_imp_1, sil_imp_2, sil_imp_3, sil_imp_4, sil_imp_5, sil_imp_6, sil_imp_7, sil_imp_8)
# sil_all_imp <-rbind(sil_all, sil_imp)
sil_imp$norm_method <- c("Linnorm", "scone", "BASiCS", "DESeq", "CPM", "TMM", "scran", "zinbwave")
sil_all$norm_method <- sil_all$method
sil_all_imp <-rbind(sil_all, sil_imp)
sil_all_imp$impu <- grepl("DrImpute$", sil_all_imp$method)
```

```{r}
# pdf("benchmark/silhouette_CEL-seq_impu.pdf")
# ggplot(sil_imp, aes(x = reorder(method, sil_width), y=sil_width, fill=reorder(method, sil_width))) + geom_boxplot() + stat_summary(fun.y=mean, geom="point",size=2) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(legend.position = "none")+ labs(x="method", y="silhouette width")
ggplot(sil_all_imp, aes(x = reorder(method, sil_width), y=sil_width, fill=reorder(method, sil_width), colour = reorder(method, sil_width))) + geom_boxplot(alpha=.5) + stat_summary(fun.y=mean, geom="point",size=2) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(legend.position = "none")+ labs(x="method", y="silhouette width")
# dev.off()
```

```{r}
ggplot(sil_all_imp, aes(x = reorder(norm_method, sil_width), y=sil_width, fill=impu, colour = impu)) + geom_boxplot(alpha=.6) + stat_summary(fun.y=mean, geom="point",size=2) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(x="method", y="silhouette width")
```



# F-statistics

```{r}
calculateF <- function(SCE, n=1){
  data <- cbind(reducedDim(SCE), colData(SCE))
  g <- lm(data[,n]~data[,"group"])
  return(anova(g)$F[1])
}
```

```{r}
f <- sapply(c(PCA1, PCA2, PCA3, PCA5, PCA6, PCA8, PCA9, PCA10, PCA11, PCA12, PCA13, PCA14), function(x){calculateF(x)}, simplify = TRUE)
f <- data.frame(f = f, method = c("raw count", "CPM", "scran", "TMM", "DESeq", "kNN", "BASiCS", "scone", "scran+DrImpute", "scran+SAVER", "Linnorm", "zinbwave"))
```

```{r}
ggplot(f, aes(x = reorder(method,f), y = f, fill=reorder(method,f))) + geom_bar(stat="identity")+ theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(legend.position = "none") + labs(x="method", y="F-statistic")
```

```{r}
f2 <- sapply(c(PCA1, PCA2, PCA3, PCA5, PCA6, PCA8, PCA9, PCA10, PCA11, PCA12, PCA13, PCA14), function(x){calculateF(x,2)}, simplify = TRUE)
f2 <- data.frame(f = f2, method = c("raw count", "CPM", "scran", "TMM", "DESeq", "kNN", "BASiCS", "scone", "scran+DrImpute", "scran+SAVER", "Linnorm", "zinbwave"))
```

```{r}
ggplot(f2, aes(x = reorder(method,f), y = f, fill=reorder(method,f))) + geom_bar(stat="identity")+ theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(legend.position = "none") + labs(x="method", y="F-statistic")
```

# save

```{r}
save.image("benchmark_mRNA.RData")
save(sce2_qc, file = "benchmark_SCE_mRNA.RData")
write.csv(f, "F/mRNA_1.csv")
write.csv(f2, "F/mRNA_2.csv")
```

