---
title: "mean silhouette width - time"
output: html_notebook
---

```{r}
library(ggplot2)
library(ggthemes)
```


#time

```{r}
meanSil_10x <- read.csv("benchmark/10x_meanSil.csv", stringsAsFactors = FALSE, row.names = 1)
meanSil_sc <- read.csv("benchmark/sc_meanSil.csv", stringsAsFactors = FALSE, row.names = 1)
meanSil_mRNA <- read.csv("benchmark/mRNA_meanSil.csv", stringsAsFactors = FALSE, row.names = 1)
meanSil_9cell <- read.csv("benchmark/9cell_meanSil.csv", stringsAsFactors = FALSE, row.names = 1)
```

```{r}
meanSil_10x$dataset <- rep("sc_10x", nrow(meanSil_10x))
meanSil_sc$dataset <- rep("sc", nrow(meanSil_sc))
meanSil_mRNA$dataset <- rep("mRNA", nrow(meanSil_mRNA))
meanSil_9cell$dataset <- rep("9cell", nrow(meanSil_9cell))
meanSil <- rbind(meanSil_10x, meanSil_sc, meanSil_mRNA, meanSil_9cell)
meanSil$dataset <- factor(meanSil$dataset, levels = c("sc", "sc_10x", "9cell", "mRNA"))
```

```{r}
kNN_row <- grepl("^kNN", meanSil$method)
meanSil[kNN_row,"method"] <- "kNN"
meanSil$method <- sub("zinbwave", "ZINB-WaVE", meanSil$method)
```

swapped x and y axis, thesis version
```{r}
library(ggrepel)

# pdf("medsil_test.pdf", width = 9)
ggplot(meanSil[meanSil$method!="scran+SAVER" & meanSil$method != "scran+DrImpute",], aes(log10(elapsed), medSil, shape= dataset, colour=method))  + geom_point(size=10, alpha=.3) + geom_text(aes(label=method))+ labs(x="log10 running time lapse", y="median silhouette width") +theme_bw()+theme(legend.position = "bottom") + scale_shape_manual(values = c(15:18))+guides(colour=FALSE)

pdf("benchmark/thesis/medsil_time_ppt.pdf", width = 13)
ggplot(meanSil[meanSil$method!="scran+SAVER" & meanSil$method != "scran+DrImpute",], aes(y=log10(elapsed), x=meanSil,  colour=method))+ geom_point( size=4, shape=20)+ geom_text_repel(aes(label=method), alpha=.7, size=5)+ labs(y="log10 running time lapse", x="mean silhouette width") +theme_bw()+ guides(colour=FALSE)+ facet_grid(.~dataset) + theme(text = element_text(size=20)) 
dev.off()
```

swapped x and y axis, paper version

```{r}
pdf("benchmark/silhouette_width_time.pdf", height=4, width=8)
library(ggrepel)
ggplot(meanSil[meanSil$method!="CPM+DrImpute"  & meanSil$dataset != "sc_10x",], aes(y=log10(elapsed), x=meanSil,  colour=method))+ geom_point( size=2, shape=20)+ geom_text_repel(aes(label=method), alpha=.7)+ labs(y="log10 running time lapse", x="mean silhouette width") +theme_bw()+ guides(colour=FALSE)+ facet_grid(.~dataset) 
dev.off()
```


# silhouette

```{r}
sil_all_10x <- read.csv("benchmark/10x_sil_all.csv", stringsAsFactors = FALSE, row.names = 1)
sil_all_sc <- read.csv("benchmark/sc_sil_all.csv", stringsAsFactors = FALSE, row.names = 1)
sil_all_mRNA <- read.csv("benchmark/mRNA_sil_all.csv", stringsAsFactors = FALSE, row.names = 1)
sil_all_9cell <- read.csv("benchmark/9cell_sil_all.csv", stringsAsFactors = FALSE, row.names = 1)
```

```{r}
sil_all_10x$dataset <- rep("sc_10x", nrow(sil_all_10x))
sil_all_sc$dataset <- rep("sc", nrow(sil_all_sc))
sil_all_mRNA$dataset <- rep("mRNA", nrow(sil_all_mRNA))
sil_all_9cell$dataset <- rep("9cell", nrow(sil_all_9cell))
sil_all <- rbind(sil_all_10x, sil_all_sc, sil_all_mRNA, sil_all_9cell)
```

```{r}
# table(sil_all$method)
kNN_row <- grepl("^kNN", sil_all$method)
sil_all[kNN_row,"method"] <- "kNN"

isImpu <- grepl("DrImpute$|SAVER$", sil_all$method)
```


## normalize the silhouette widths by raw count

```{r}
head(sil_all)
meanSil_mRNA_raw <- meanSil_mRNA[meanSil_mRNA$method == "raw count", "meanSil"]
meanSil_9cell_raw <- meanSil_9cell[meanSil_9cell$method == "raw count", "meanSil"]
meanSil_sc_raw <- meanSil_sc[meanSil_sc$method == "raw count", "meanSil"]
meanSil_10x_raw <- meanSil_10x[meanSil_10x$method == "raw count", "meanSil"]

```

```{r}
sil_9cell_norm <- cbind(sil_all_9cell, sil_width_norm = sil_all_9cell$sil_width - meanSil_9cell_raw)
sil_mRNA_norm <- cbind(sil_all_mRNA, sil_width_norm = sil_all_mRNA$sil_width - meanSil_mRNA_raw)
sil_sc_norm <- cbind(sil_all_sc, sil_width_norm = sil_all_sc$sil_width - meanSil_sc_raw)
sil_10x_norm <- cbind(sil_all_10x, sil_width_norm = sil_all_10x$sil_width - meanSil_10x_raw)
sil_norm <- rbind(sil_9cell_norm, sil_mRNA_norm, sil_sc_norm, sil_10x_norm)

isImpu_norm <- grepl("DrImpute$|SAVER$", sil_norm$method)

kNN_row <- grepl("^kNN", sil_norm$method)
sil_norm[kNN_row,"method"] <- "kNN"
sil_norm$method <- sub("zinbwave", "ZINB-WaVE", sil_norm$method)

```

thesis version
```{r}
order_thesis <- unique(reorder(sil_norm$method, sil_norm$sil_width_norm))
sil_norm$method <- factor(sil_norm$method, levels(order_thesis))

pdf("benchmark/thesis/silhouette_norm.pdf", height=4, width=8)
ggplot(sil_norm[sil_norm$method != "raw count" & sil_norm$method != "ZINB-WaVE W"& isImpu_norm == FALSE,], aes(x =method, y=sil_width_norm, fill=dataset,colour=dataset))+geom_boxplot(alpha=.5)   + theme_bw() +scale_fill_calc()+scale_colour_calc() + theme(axis.text.x = element_text(angle = 30, hjust = 1),text = element_text(size=14))+ labs(x="method", y="normalized silhouette width") 
dev.off()
```




paper version with zinbwave W value

```{r}
sil_norm_cel <- sil_norm[sil_norm$dataset!="sc_10x"& sil_norm$method != "CPM_DrImpute",]
order <- unique(reorder(sil_norm_cel$method, sil_norm_cel$sil_width_norm))
sil_norm_cel$method <- factor(sil_norm_cel$method, levels(order))




pdf("benchmark/silhouette_CEL-seq2_norm.pdf", height=4, width=8)
ggplot(sil_norm_cel[sil_norm_cel$method != "raw count",], aes(x = method, y=sil_width_norm, fill=dataset,colour=dataset))+geom_boxplot(alpha=.5)   + theme_bw() +scale_fill_calc()+scale_colour_calc() + theme(axis.text.x = element_text(angle = 30, hjust = 1),text = element_text(size=20))+ labs(x="method", y="normalized silhouette width") 

dev.off()
```

paper version, no W

```{r}
pdf("benchmark/silhouette_no_W_norm.pdf", height=4, width=8)
ggplot(sil_norm_cel[sil_norm_cel$method != "raw count" & sil_norm_cel$method != "ZINB-WaVE W",], aes(x = method, y=sil_width_norm, fill=dataset,colour=dataset))+geom_boxplot(alpha=.5)   + theme_bw() +scale_fill_calc()+scale_colour_calc() + theme(axis.text.x = element_text(angle = 30, hjust = 1),text = element_text(size=20), axis.title.y = element_text(size=16))+ labs(x="method", y="normalized silhouette width") 

dev.off()
```

paper version, not normalized

```{r}
sil_all_cel <- sil_all[sil_all$dataset != "sc_10x"& sil_all$method != "CPM_DrImpute"& sil_all$method != "zinbwave W",]
sil_all_cel$method <- sub("zinbwave", "ZINB-WaVE", sil_all_cel$method)
sil_all_cel$method <- factor(sil_all_cel$method, levels(order))

pdf("benchmark/silhouette_raw.pdf", height=4, width=8)
ggplot(sil_all_cel[sil_all_cel$method!="ZINB-WaVE W",], aes(x = method, y=sil_width, fill=dataset,colour=dataset))+geom_boxplot(alpha=.5)   + theme_bw() +scale_fill_calc()+scale_colour_calc() + theme(axis.text.x = element_text(angle = 30, hjust = 1),text = element_text(size=20))+ labs(x="method", y="silhouette width") 

dev.off()
```

# correlation
```{r}
cor_df <- read.csv("benchmark/cor_df.csv", stringsAsFactors = FALSE, row.names = 1)
cor_df$method <- sub("linnorm", "Linnorm", cor_df$method)
# head(cor_df)
```

```{r}
# table(cor_df$method)


# intersect(order, cor_df$method)
# order
cor_df$method <- factor(cor_df$method, intersect(levels(order), cor_df$method))
```

paper version correlation

```{r}
pdf("benchmark/cor_df.pdf", height = 4, width = 8)
ggplot(cor_df[cor_df$method != "raw count",], aes(method, cor_val, colour=factor(mRNA_amount),fill=factor(mRNA_amount))) + geom_boxplot(alpha=.5) + theme_bw() + scale_fill_brewer(name="mRNA amount", palette = "Dark2")+scale_colour_brewer(name="mRNA amount", palette = "Dark2")+ theme(axis.text.x = element_text(angle = 30, hjust = 1),text = element_text(size=20)) 
dev.off()
```

thesis version
```{r}
cor_df$method <- factor(cor_df$method, intersect(levels(order_thesis), cor_df$method))

pdf("benchmark/thesis/cor_df.pdf", height = 4, width = 8)
ggplot(cor_df[cor_df$method != "raw count",], aes(method, cor_val, colour=factor(mRNA_amount),fill=factor(mRNA_amount))) + geom_boxplot(alpha=.5) + theme_bw() + scale_fill_brewer(name="mRNA amount", palette = "Dark2")+scale_colour_brewer(name="mRNA amount", palette = "Dark2")+ theme(axis.text.x = element_text(angle = 30, hjust = 1),text = element_text(size=14)) 
dev.off()

```



# 9 cell silhouette width

```{r}
sil_all_9cell1 <- read.csv("benchmark/9cell1_sil_all.csv", stringsAsFactors = FALSE, row.names = 1)
sil_all_9cell2 <- read.csv("benchmark/9cell2_sil_all.csv", stringsAsFactors = FALSE, row.names = 1)
sil_all_9cell4 <- read.csv("benchmark/9cell4_sil_all.csv", stringsAsFactors = FALSE, row.names = 1)
sil_all_9cell3 <- sil_all_9cell

sil_all_9cell1$dataset <- rep("9cell1", nrow(sil_all_9cell1))
sil_all_9cell2$dataset <- rep("9cell2", nrow(sil_all_9cell2))
sil_all_9cell3$dataset <- rep("9cell3", nrow(sil_all_9cell3))
sil_all_9cell4$dataset <- rep("9cell4", nrow(sil_all_9cell4))

sil_9cell <- rbind(sil_all_9cell1, sil_all_9cell2, sil_all_9cell3, sil_all_9cell4)
```

```{r}
table(sil_9cell$method)
sil_9cell <- sil_9cell[sil_9cell$method !="CPM_DrImpute" &sil_9cell$method !="zinbwave W" ,]
kNN_row <- grepl("^kNN", sil_9cell$method)
sil_9cell[kNN_row,"method"] <- "kNN"
sil_9cell$method <- sub("zinbwave", "ZINB-WaVE", sil_9cell$method)
# sil_9cell$method <- factor(sil_9cell$method, intersect(levels(order), sil_9cell$method))
```

```{r}
pdf("benchmark/thesis/silhouette_width_9cell.pdf", height=4, width=8)
ggplot(sil_9cell, aes(x = reorder(method, sil_width), y=sil_width, fill=dataset,colour=dataset))+geom_boxplot(alpha=.5)   + theme_bw() +scale_fill_brewer(palette = "Dark2")+scale_colour_brewer(palette = "Dark2") + theme(axis.text.x = element_text(angle = 30, hjust = 1),text = element_text(size=20))+ labs(x="method", y="silhouette width") 
dev.off()
```

# Imputation

scran-related methods

```{r}
is_scran <- grepl("^scran", sil_all$method)
table(is_scran)

ggplot(sil_all[is_scran,], aes(x=dataset, y=sil_width, fill=method))+geom_violin()+scale_fill_brewer(palette="Accent")+theme_bw()
```

```{r}
library(ggridges)
ggplot(sil_all[is_scran,], aes(y=method, x=sil_width, fill=method))+scale_fill_brewer(palette="Accent")+geom_density_ridges(scale=3, alpha=.7)+ facet_grid(dataset~.)+theme_ridges()
```

# F-statistic

```{r}
f_sc <- read.csv("F/sc_1.csv")
f_9cell <- read.csv("F/9cell3_1.csv")
f_mRNA <- read.csv("F/mRNA_1.csv")
f_10x <- read.csv("F/sc_10x_1.csv")

f_sc$dataset <- "sc"
f_9cell$dataset <- "9cell"
f_mRNA$dataset <- "mRNA"
f_10x$dataset <- "sc_10x"
```

```{r}
f2_sc <- read.csv("F/sc_2.csv")
f2_9cell <- read.csv("F/9cell3_2.csv")
f2_mRNA <- read.csv("F/mRNA_2.csv")
f2_10x <- read.csv("F/sc_10x_2.csv")

# f2_sc$dataset <- "sc"
# f2_9cell$dataset <- "9cell"
# f2_mRNA$dataset <- "mRNA"
# f2_10x$dataset <- "sc_10x"
```

```{r}
f_sc$f2 <- f2_sc$f
f_9cell$f2 <- f2_9cell$f
f_mRNA$f2 <- f2_mRNA$f
f_10x$f2 <- f2_10x$f
```

```{r}
f_sc$f_all <- f_sc$f + f_sc$f2
f_9cell$f_all <- f_9cell$f + f_9cell$f2
f_mRNA$f_all <- f_mRNA$f + f_mRNA$f2
f_10x$f_all <- f_10x$f + f_10x$f2
```

<!--
normalization by dividing median F value
```{r}
f_sc$f_norm <- f_sc$f / median(f_sc$f)
f_9cell$f_norm <- f_9cell$f / median(f_9cell$f)
f_mRNA$f_norm <- f_mRNA$f / median(f_mRNA$f)
f_10x$f_norm <- f_10x$f / median(f_10x$f)
f2_sc$f_norm <- f2_sc$f / median(f2_sc$f)
f2_9cell$f_norm <- f2_9cell$f / median(f2_9cell$f)
f2_mRNA$f_norm <- f2_mRNA$f / median(f2_mRNA$f)
f2_10x$f_norm <- f2_10x$f / median(f2_10x$f)
f_sc$f_all_norm <- f_sc$f_all / median(f_sc$f_all)
f_9cell$f_all_norm <- f_9cell$f_all / median(f_9cell$f_all)
f_mRNA$f_all_norm <- f_mRNA$f_all / median(f_mRNA$f_all)
f_10x$f_all_norm <- f_10x$f_all / median(f_10x$f_all)
```

-->

```{r}
f <- rbind(f_sc, f_10x, f_9cell, f_mRNA)
f2 <- rbind(f2_sc, f2_10x, f2_9cell, f2_mRNA)
# f$PC <- "1"
# f2$PC <- "2"
# f_all <- rbind(f,f2)
# f$f_all <- f$f + f2$f
# f$f_norm_all <- f$f_norm + f2$f_norm
# f$f2 <- f2$f

# f$f2_norm <- f2$f_norm
f$method <- sub("zinbwave", "ZINB-WaVE", f$method)
```

normalization
```{r}
for(i in unique(f$dataset)){
  f[f$dataset == i, "f_all_norm"] <- (f[f$dataset==i, "f_all"] - f[f$dataset==i & f$method=="raw count", "f_all"]) / f[f$dataset==i & f$method=="raw count", "f_all"]
}
```


bar plot
```{r}

f$method <- factor(f$method, intersect(levels(order_thesis), f$method))

ggplot(f[f$method != "raw count" & f$method != "scran+SAVER" & f$method != "scran+DrImpute",], aes(x=method, y=f_all_norm, fill=dataset, colour=dataset))+geom_bar(stat = "identity", position=position_dodge(), alpha=.7)+ theme_bw() +scale_fill_calc() +scale_colour_calc()+ theme(axis.text.x = element_text(angle = 30, hjust = 1),text = element_text(size=14))+ labs(x="method", y="normalized F-statistic") 

```

```{r}
plot_sil <- ggplot(sil_norm[sil_norm$method != "raw count" & sil_norm$method != "ZINB-WaVE W"& isImpu_norm == FALSE,], aes(x =method, y=sil_width_norm, fill=dataset,colour=dataset))+geom_boxplot(alpha=.5)   + theme_bw() +scale_fill_calc()+scale_colour_calc() + theme(axis.text.x = element_text(angle = 30, hjust = 1),text = element_text(size=14), legend.position="none")+ labs(x="method", y="normalized silhouette width", title="A. Normalized silhouette width") 

f$method <- factor(f$method, intersect(levels(order_thesis), f$method))
plot_f <- ggplot(f[f$method != "raw count" & f$method != "scran+SAVER" & f$method != "scran+DrImpute",], aes(x=method, y=f_all_norm, fill=dataset, colour=dataset))+geom_bar(stat = "identity", position=position_dodge(), alpha=.7)+ theme_bw() +scale_fill_calc() +scale_colour_calc()+ theme(axis.text.x = element_text(angle = 30, hjust = 1),text = element_text(size=14), legend.position="bottom")+ labs(x="method", y="normalized F-statistic", title="B. Normalized F-statistic") 

pdf("benchmark/thesis/sil+F_ppt.pdf", width=12)
multiplot(plot_sil, plot_f)
dev.off()
```

```{r}
plot_f1 <- ggplot(f[f$method != "raw count" & f$method != "scran+SAVER" & f$method != "scran+DrImpute",], aes(x=method, y=f_norm, fill=dataset, colour=dataset))+geom_bar(stat = "identity", position=position_dodge(), alpha=.7)+ theme_bw() +scale_fill_calc() +scale_colour_calc()+ theme(axis.text.x = element_text(angle = 30, hjust = 1),text = element_text(size=14), legend.position="bottom")+ labs(x="method", y="normalized F-statistic", title="A. Normalized F-statistic of PC1") 

plot_f2 <- ggplot(f[f$method != "raw count" & f$method != "scran+SAVER" & f$method != "scran+DrImpute",], aes(x=method, y=f2_norm, fill=dataset, colour=dataset))+geom_bar(stat = "identity", position=position_dodge(), alpha=.7)+ theme_bw() +scale_fill_calc() +scale_colour_calc()+ theme(axis.text.x = element_text(angle = 30, hjust = 1),text = element_text(size=14), legend.position="bottom")+ labs(x="method", y="normalized F-statistic", title="B. Normalized F-statistic of PC2")

pdf("benchmark/thesis/F1+F2.pdf")
multiplot(plot_f1, plot_f2)
dev.off()
```

