---
title: "10x benchmarking"
output: html_notebook
---

# Preparation

Load some packages

```{r message=FALSE}
library(scran)
library(scater)
library(ggplot2)
library(edgeR)
library(DESeq)
library(ggridges)
```

Load R data

```{r}
# load("SCE_sc.RData")
load("SCE_sc_newQC.RData")
```

Assign cell line information and check

```{r}
# sce10x_qc$cell_line <- sce10x_qc_sel$cell_line
# sce4_qc$cell_line <- sce4_qc_sel$cell_line
# sce701_qc$cell_line <- sce701_qc_sel$cell_line

table(sce4_qc$cell_line)
table(sce10x_qc$cell_line)
table(sce701_qc$cell_line)
```

Function for calculate silhouette width of PCA analysis, group samples according to their cell type

```{r}
library(cluster)

calcSilhouette <- function(SCE, name){
  sil <- silhouette(as.numeric(factor(SCE$cell_line)), dist(reducedDim(SCE)))
  sil <- data.frame(sil[1:nrow(sil),])
  method <- rep(name, nrow(sil))
  sil <- cbind(sil, method)
}
```

Function for subsample of a specific cell line (reduce the number of cells of a specific cell line)
useful for test whether algorithms are robust and preserve rare cell type

```{r}
subsample_cell <- function(SCE, cell_line, proportion){
  cells <- which(SCE$cell_line == cell_line)
  delete_n <- length(cells) * (1 - proportion) # number of cells to be deleted
  delete_cell <- sample(cells, delete_n)
  SCE_return <- SCE[,-delete_cell]
  assays(SCE_return) <- list(counts = counts(SCE_return)) # delete useless assays to reduce its size
  return (SCE_return)
}
```


# benchmarking for 10x data set

## raw log count

```{r}
time_raw <- system.time(
logcounts(sce10x_qc) <- log2(counts(sce10x_qc) + 1)
)
# cor_df_raw_10x = cor_sc_within(sce10x_qc, "logcounts")
# ggplot(data = cor_df_raw_10x, aes(x = factor(cell_line), y = cor_val, fill = factor(cell_line))) + geom_violin()
```

## log CPM

```{r}
time_CPM <- system.time({
assay(sce10x_qc, "norm_cpm") = log2(cpm(counts(sce10x_qc)) + 1)
})
# cor_df_cpm_10x <- cor_sc_within(sce10x_qc, "norm_cpm")
# ggplot(data = cor_df_cpm_10x, aes(x = factor(cell_line), y = cor_val, fill = factor(cell_line))) + geom_violin()
```

## scran

```{r}
time_scran <- system.time({
sce10x_qc <- computeSumFactors(sce10x_qc)
sce10x_qc <- normalize(sce10x_qc)
assay(sce10x_qc, "scran") = logcounts(sce10x_qc)
})
# cor_df_scran_10x <- cor_sc_within(sce10x_qc, "scran")
# ggplot(data = cor_df_scran_10x, aes(x = factor(cell_line), y = cor_val, fill = factor(cell_line))) + geom_violin()
```

## TMM

```{r}
time_TMM <- system.time({
sizeFactors(sce10x_qc) <- calcNormFactors(counts(sce10x_qc), method = "TMM")
sce10x_qc <- normalize(sce10x_qc)
assay(sce10x_qc, "TMM") <- logcounts(sce10x_qc)
})
# cor_df_TMM_10x <- cor_sc_within(sce10x_qc, "TMM")
# ggplot(data = cor_df_TMM_10x, aes(x = factor(cell_line), y = cor_val, fill = factor(cell_line))) + geom_violin()

```

## DESeq

```{r}
time_DESeq <- system.time({
sizeFactors(sce10x_qc) <- estimateSizeFactorsForMatrix(counts(sce10x_qc))
sce10x_qc <- normalize(sce10x_qc)
assay(sce10x_qc, "DESeq") <- logcounts(sce10x_qc)
})
# cor_df_DESeq_10x <- cor_sc_within(sce10x_qc, "DESeq")
# ggplot(data = cor_df_DESeq_10x, aes(x = factor(cell_line), y = cor_val, fill = factor(cell_line))) + geom_violin()
```

## SCnorm

```{r}
# load("SCnorm_out_10x.RData")
# ```
# 
# ```{r}
# assay(sce10x_qc, "SCnorm") = log2(metadata(SCnorm_out_10x)$NormalizedData + 1)
# cor_df_SCnorm_10x <- cor_sc_within(sce10x_qc, "SCnorm")
# ggplot(data = cor_df_SCnorm_10x, aes(x = factor(cell_line), y = cor_val, fill = factor(cell_line))) + geom_violin()
```
```{r}
library(SCnorm)
#time1 <- Sys.time()
time_SCnorm <- system.time({
  SCnorm_out <- SCnorm(Data = counts(sce10x_qc), Conditions = as.numeric(factor(sce10x_qc$cell_line)), PrintProgressPlots = T)
  assay(sce10x_qc, "SCnorm") = log2(metadata(SCnorm_out)$NormalizedData + 1)
 
})
save(SCnorm_out, file = "SCnorm_out_new.RData")

detach("package:SCnorm", unload=TRUE) 
```



## knn

```{r}
source("/wehisan/home/allstaff/d/dong.x/mixture/knn-smoothing-master/knn_smooth.R")
time_knn63 <- system.time({
sce4_smoothed = knn_smoother(counts(sce10x_qc), k = 63)
assay(sce10x_qc, "knn63") = log2(sce4_smoothed + 1)
})
```

## `scone` normalization

```{r}
library(scone)

expr <- counts(sce10x_qc)
bio <- factor(sce10x_qc$cell_line)


#Read in housekeeping gene list and convert gene symbol to ensembl gene ID using biomaRt.


hkgene <- read.csv("h-scHKgenes.csv", stringsAsFactors = FALSE, col.names = 1)
library(biomaRt)
ensembl = useMart("ensembl")
ensembl = useDataset("hsapiens_gene_ensembl", mart = ensembl)
hk <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"), filters = "hgnc_symbol", mart = ensembl, values = hkgene$X1)


#Use housekeeping genes as negetive control
#time1 <- Sys.time()
time_scone <- system.time({
  negcon <- intersect(rownames(expr), hk$ensembl_gene_id)
  # creating a SconeExperiment Object
  my_scone <- SconeExperiment(expr, bio = bio, negcon_ruv = rownames(expr) %in% negcon)
  
  scaling=list(none=identity, # Identity - do nothing
               
               # eff = EFF_FN, # User-defined function
               
               sum = SUM_FN, # SCONE library wrappers...
               tmm = TMM_FN, 
               uq = UQ_FN,
               fq = FQT_FN,
               deseq = DESEQ_FN)
  
  BiocParallel::register(
    BiocParallel::SerialParam()
  ) # Register BiocParallel Serial Execution
  
  my_scone <- scone(my_scone,
                    #imputation = imputation, impute_args = impute_args,
                    scaling=scaling,
                    run=TRUE,
                    eval_kclust = 2:6,stratified_pam = TRUE,
                    k_qc=0, k_ruv = 3,
                    adjust_bio="no",
                    return_norm = "in_memory",
                    zero = "postadjust")
  # NO RUV
  
  
  # load("my_scone.RData")
  # library(scone)
  
  scores <- get_scores(my_scone)
  score_ranks <- get_score_ranks(my_scone)
  out_norm = get_normalized(my_scone,method = rownames(get_params(my_scone))[1])
  
  assay(sce10x_qc, "scone") = log2(out_norm + 1)
})
# time2 <- Sys.time()
# time_scone <- c("scone", time2 - time1)
# cor_df_scone <- cor_sc_within(sce4_qc, "scone")
# ggplot(data = cor_df_scone, aes(x = factor(cell_line), y = cor_val, fill = factor(cell_line))) + geom_violin()
detach("package:scone", unload=TRUE) 
detach("package:biomaRt", unload=TRUE) 
```
## scran + DrImpute

```{r}
library(DrImpute)
time_DrImpute_scran <-system.time({
imp_10x <- DrImpute(assay(sce10x_qc, "scran"))
assay(sce10x_qc, "scran_DrImpute") <- imp_10x
})
detach("package:DrImpute", unload=TRUE) 
# cor_df_DrImpute_10x = cor_sc_within(sce10x_qc, "scran_DrImpute")
# ggplot(data = cor_df_DrImpute_10x, aes(x = factor(cell_line), y = cor_val, fill = factor(cell_line))) + geom_violin()
```

<!--
## total count
```{r}
library(DrImpute)
sf <- apply(counts(sce10x_qc), 2, mean)
tcount <- t(t(counts(sce10x_qc)) / sf)
tcount <- log(tcount + 1)
impcount <- DrImpute(tcount)
assay(sce10x_qc, "totalcount") <- tcount
assay(sce10x_qc, "totalcount_DrImpute") <- impcount
#cor_df_totalcount = cor_sc_within(sce4_qc, "totalcount")
#cor_df_DrImputeTotalcount = cor_sc_within(sce4_qc, "totalcount_DrImpute")
#ggplot(data = cor_df_totalcount, aes(x = factor(cell_line), y = cor_val, fill = factor(cell_line))) + geom_violin()+ ggtitle("total count")
#ggplot(data = cor_df_DrImputeTotalcount, aes(x = factor(cell_line), y = cor_val, fill = factor(cell_line))) + geom_violin()+ ggtitle("total count+DrImpute")
#testp1 <- plotPCASCE(sce4_qc, exprs_values = "totalcount_DrImpute", colour_by = "cell_line") + ggtitle("total count+DrImpute")
#testp2 <-plotPCASCE(sce4_qc, exprs_values = "totalcount", colour_by = "cell_line") + ggtitle("total count")
#multiplot(testp2 ,testp1)
```
-->

## SAVER + scran

```{r}
library(SAVER)
library(doParallel)
time_SAVER <- system.time({
  sce10x_qc <- computeSumFactors(sce10x_qc)
  registerDoParallel(cores = 32)
  saver_out <- saver(counts(sce10x_qc), size.factor = sizeFactors(sce10x_qc))
  #load("saver_out.RData")
  assay(sce10x_qc, "SAVER_scran") <- log2(saver_out$estimate + 1)
})
# cor_df_SAVER = cor_sc_within(sce4_qc, "SAVER_scran")
# ggplot(data = cor_df_SAVER, aes(x = factor(cell_line), y = cor_val, fill = factor(cell_line))) + geom_violin()
```

## Linnorm

```{r}
library(Linnorm)
time_Linnorm <- system.time({
linnorm_normalized_10x <- Linnorm(counts(sce10x_qc))
assay(sce10x_qc, "linnorm") <- linnorm_normalized_10x
# cor_df_linnorm_10x = cor_sc_within(sce10x_qc, "linnorm")
})
# ggplot(data = cor_df_linnorm_10x, aes(x = factor(cell_line), y = cor_val, fill = factor(cell_line))) + geom_violin()
detach("package:Linnorm", unload=TRUE) 
```

## zinbwave

```{r}
load("zinbwave_sc_10x.RData")
assay(sce10x_qc, "zinbwave") <- assay(se_norm, "normalizedValues")

```



# results intergration

## PCA plot visualization

```{r}
PCA1 <- runPCA(sce10x_qc, exprs_values = "counts", ncomponents = 2)
PCA2 <- runPCA(sce10x_qc, exprs_values = "norm_cpm", ncomponents = 2)
PCA3 <- runPCA(sce10x_qc, exprs_values = "scran", ncomponents = 2) 
PCA4 <- runPCA(sce10x_qc, exprs_values = "SCnorm", ncomponents = 2) 
PCA5 <- runPCA(sce10x_qc, exprs_values = "TMM", ncomponents = 2) 
PCA6 <- runPCA(sce10x_qc, exprs_values = "DESeq", ncomponents = 2)
# PCA7 <- runPCA(sce10x_qc, exprs_values = "knn15", ncomponents = 2)
PCA8 <- runPCA(sce10x_qc, exprs_values = "knn63", ncomponents = 2) 
PCA9 <- runPCA(sce10x_qc, exprs_values = "scone", ncomponents = 2) 
PCA10 <- runPCA(sce10x_qc, exprs_values = "scran_DrImpute", ncomponents = 2) 
PCA11 <- runPCA(sce10x_qc, exprs_values = "SAVER_scran", ncomponents = 2) 
PCA12 <- runPCA(sce10x_qc, exprs_values = "linnorm", ncomponents = 2) 
# PCA13 <- runPCA(sce10x_qc, exprs_values = "totalcount", ncomponents = 2) 
# PCA14 <- runPCA(sce10x_qc, exprs_values = "totalcount_DrImpute", ncomponents = 2) 
PCA13 <- runPCA(sce10x_qc, exprs_values = "zinbwave", ncomponents = 2) 


p1 <- plotPCASCE(PCA1, colour_by = "cell_line", exprs_values = "counts", ncomponents = 2) + ggtitle("raw counts")+ theme(legend.position="none")+ scale_fill_manual(values = c("#FF0000", "#00FF00", "#0000FF"), limits = c("H1975", "H2228", "HCC827"))
p2 <- plotPCASCE(PCA2, colour_by = "cell_line", exprs_values = "norm_cpm", ncomponents = 2) + ggtitle("CPM")+ theme(legend.position="none")+ scale_fill_manual(values = c("#FF0000", "#00FF00", "#0000FF"), limits = c("H1975", "H2228", "HCC827"))
p3 <- plotPCASCE(PCA3, colour_by = "cell_line", exprs_values = "scran", ncomponents = 2) + ggtitle("scran")+ theme(legend.position="none")+ scale_fill_manual(values = c("#FF0000", "#00FF00", "#0000FF"), limits = c("H1975", "H2228", "HCC827"))
p4 <- plotPCASCE(PCA4, colour_by = "cell_line", exprs_values = "SCnorm", ncomponents = 2) + ggtitle("SCnorm")+ theme(legend.position="none")+ scale_fill_manual(values = c("#FF0000", "#00FF00", "#0000FF"), limits = c("H1975", "H2228", "HCC827"))
p5 <- plotPCASCE(PCA5, colour_by = "cell_line", exprs_values = "TMM", ncomponents = 2) + ggtitle("TMM")+ theme(legend.position="none")+ scale_fill_manual(values = c("#FF0000", "#00FF00", "#0000FF"), limits = c("H1975", "H2228", "HCC827"))
p6 <- plotPCASCE(PCA6, colour_by = "cell_line", exprs_values = "DESeq", ncomponents = 2) + ggtitle("DESeq")+ theme(legend.position="none")+ scale_fill_manual(values = c("#FF0000", "#00FF00", "#0000FF"), limits = c("H1975", "H2228", "HCC827"))
# p7_10x <- plotPCASCE(PCA7_10x, colour_by = "cell_line", exprs_values = "knn", ncomponents = 2) + ggtitle("knn (k=15)")+ theme(legend.position="none")
p8 <- plotPCASCE(PCA8, colour_by = "cell_line", exprs_values = "knn63", ncomponents = 2) + ggtitle("knn (k=63)")+ theme(legend.position="none")+ scale_fill_manual(values = c("#FF0000", "#00FF00", "#0000FF"), limits = c("H1975", "H2228", "HCC827"))
p9 <- plotPCASCE(PCA9, colour_by = "cell_line", exprs_values = "scone", ncomponents = 2) + ggtitle("scone")+ theme(legend.position="none")+ scale_fill_manual(values = c("#FF0000", "#00FF00", "#0000FF"), limits = c("H1975", "H2228", "HCC827"))
p10 <- plotPCASCE(PCA10, colour_by = "cell_line", exprs_values = "scran_DrImpute", ncomponents = 2) + ggtitle("scran+DrImpute")+ theme(legend.position="none")+ scale_fill_manual(values = c("#FF0000", "#00FF00", "#0000FF"), limits = c("H1975", "H2228", "HCC827"))
p11 <- plotPCASCE(PCA11, colour_by = "cell_line", exprs_values = "SAVER_scran", ncomponents = 2) + ggtitle("SAVER+scran")+ theme(legend.position="none")+ scale_fill_manual(values = c("#FF0000", "#00FF00", "#0000FF"), limits = c("H1975", "H2228", "HCC827"))
p12 <- plotPCASCE(PCA12, colour_by = "cell_line", exprs_values = "linnorm", ncomponents = 2) + ggtitle("Linnorm")+ theme(legend.position="none")+ scale_fill_manual(values = c("#FF0000", "#00FF00", "#0000FF"), limits = c("H1975", "H2228", "HCC827"))
p13 <- plotPCASCE(PCA13, colour_by = "cell_line", exprs_values = "zinbwave", ncomponents = 2) + ggtitle("zinbwave")+ theme(legend.position="none")+ scale_fill_manual(values = c("#FF0000", "#00FF00", "#0000FF"), limits = c("H1975", "H2228", "HCC827"))
# p13 <- plotPCASCE(PCA13, colour_by = "cell_line", exprs_values = "totalcount", ncomponents = 2) + ggtitle("total count")+ theme(legend.position="none")
# p14 <- plotPCASCE(PCA14, colour_by = "cell_line", exprs_values = "totalcount_DrImpute", ncomponents = 2) + ggtitle("total count+DrImpute")+ theme(legend.position="none")

pdf("benchmark/thesis/PCA_sc_10x.pdf", height = 9)
multiplot(p1, p2, p3, p4, p5, p6,  p8, p9,  p12, p13, cols = 3)
dev.off()
```

```{r}
sil1 <- calcSilhouette(PCA1, "raw count")
sil2 <- calcSilhouette(PCA2, "CPM")
sil3 <- calcSilhouette(PCA3, "scran")
sil4 <- calcSilhouette(PCA4, "SCnorm")
sil5 <- calcSilhouette(PCA5, "TMM")
sil6 <- calcSilhouette(PCA6, "DESeq")
# sil7 <- calcSilhouette(PCA7, "kNN (k=15)")
sil8 <- calcSilhouette(PCA8, "kNN (k=63)")
sil9 <- calcSilhouette(PCA9, "scone")
sil10 <- calcSilhouette(PCA10, "scran+DrImpute")
sil11 <- calcSilhouette(PCA11, "scran+SAVER")
sil12 <- calcSilhouette(PCA12, "Linnorm")
sil13 <- calcSilhouette(PCA13, "zinbwave")
# sil13 <- calcSilhouette(PCA13, "total count")
# sil14 <- calcSilhouette(PCA14, "total count+DrImpute")
sil14 <- calcSilhouette(se_norm, "zinbwave W")

sil_all <- rbind(sil1, sil2, sil3, sil4, sil5, sil6,  sil8, sil9, sil10, sil11, sil12, sil13, sil14)
# pdf("benchmark/silhouette_10x.pdf")
ggplot(sil_all, aes(x = reorder(method, sil_width), y=sil_width, fill = reorder(method, sil_width))) + geom_boxplot() + stat_summary(fun.y=mean, geom="point",size=2) + theme_bw() + ggtitle("Silhouette width of 10x data") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(legend.position = "none")
# dev.off()
```

```{r}
# sil_comb = rbind(cbind(sil_all, platform = rep("CEL-seq2", nrow(sil_all))), cbind(sil_all_10x, platform = rep("10x", nrow(sil_all_10x))))
# #pdf("benchmark/silhouette.pdf")
# ggplot(sil_comb, aes(x = method, y = sil_width, fill = platform)) + geom_boxplot() + theme_bw()+ theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(legend.position = "none")
# #dev.off()
# ```
# ```{r}
# #pdf("benchmark/silhouette_joy.pdf")
# ggplot(sil_comb, aes(x = sil_width, y = method, fill = platform)) + geom_density_ridges(alpha = .4) + theme_ridges()
# #+geom_vline(xintercept = .8, alpha = .5, colour = "blue")+geom_vline(xintercept = .9, alpha = .5, colour = "blue")
# #dev.off()
```

### timer

box plot coloured by time elapsed
```{r}
library(viridis)
time_all <- rbind(time_CPM, time_DESeq, time_DrImpute_scran,  time_knn63,  time_raw, time_SCnorm, time_scone, time_scran, time_TMM, time_Linnorm, time_SAVER, time_zinbwave)
time_all <- data.frame(time_all)
time_all$method <- c("CPM", "DESeq", "scran+DrImpute",  "kNN (k=63)", "raw count", "SCnorm", "scone", "scran", "TMM", "Linnorm", "scran+SAVER", "zinbwave")
time_all_sil <- merge(sil_all, time_all, by = "method")
# pdf("benchmark/silhouette_time_CEL-seq.pdf")
# ggplot(time_all_sil, aes(x = reorder(method, sil_width), y=sil_width, fill=log10(elapsed))) + geom_boxplot() + stat_summary(fun.y=mean, geom="point",size=2) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + scale_fill_viridis()
# dev.off()
#dim(time_all_sil)
```



```{r}
meanSil <- sapply(unique(time_all_sil$method), function(x){
  mean(time_all_sil[time_all_sil$method == x, "sil_width"])
}, simplify = TRUE)
medSil <- sapply(unique(time_all_sil$method), function(x){
  median(time_all_sil[time_all_sil$method == x, "sil_width"])
}, simplify = TRUE)
meanSil <- data.frame(meanSil, medSil, method = unique(time_all_sil$method))
meanSil <- merge(meanSil, time_all, by = "method")

# head(meanSil)

# pdf("benchmark/medsiltime_10x.pdf", height=4)
ggplot(meanSil, aes(log10(elapsed), medSil, colour= method))  + geom_text(aes(label=method)) + labs(x="log10 running time lapse", y="median silhouette width") +theme_bw()+ theme(legend.position = "none")
# dev.off()

write.csv(meanSil, "benchmark/10x_meanSil.csv")
```

# imputation

```{r impute}
library(DrImpute)
#Linnorm
imp <- DrImpute(assay(sce10x_qc, "linnorm"))
assay(sce10x_qc, "linnorm_DrImpute") <- imp
#scone
imp <- DrImpute(assay(sce10x_qc, "scone"))
assay(sce10x_qc, "scone_DrImpute") <- imp
#SCnorm
imp <- DrImpute(assay(sce10x_qc, "SCnorm"))
assay(sce10x_qc, "SCnorm_DrImpute") <- imp
#DESeq
imp <- DrImpute(assay(sce10x_qc, "DESeq"))
assay(sce10x_qc, "DESeq_DrImpute") <- imp
#CPM
imp <- DrImpute(assay(sce10x_qc, "norm_cpm"))
assay(sce10x_qc, "CPM_DrImpute") <- imp
#TMM
imp <- DrImpute(assay(sce10x_qc, "TMM"))
assay(sce10x_qc, "TMM_DrImpute") <- imp
```

```{r PCA_impute}

PCA_imp1 <- runPCA(sce10x_qc, exprs_values = "linnorm_DrImpute")
PCA_imp2 <- runPCA(sce10x_qc, exprs_values = "scone_DrImpute")
PCA_imp3 <- runPCA(sce10x_qc, exprs_values = "SCnorm_DrImpute")
# PCA_imp10x <- runPCA(sce10x_qc, exprs_values = "BASiCS_DrImpute")
PCA_imp5 <- runPCA(sce10x_qc, exprs_values = "DESeq_DrImpute")
PCA_imp6 <- runPCA(sce10x_qc, exprs_values = "CPM_DrImpute")
PCA_imp7 <- runPCA(sce10x_qc, exprs_values = "TMM_DrImpute")

p_imp1 <- plotPCASCE(PCA_imp1, colour_by = "cell_line", exprs_values = "linnorm_DrImpute", ncomponents = 2) + ggtitle("linnorm_DrImpute")+ theme(legend.position="none")
p_imp2 <- plotPCASCE(PCA_imp2, colour_by = "cell_line", exprs_values = "scone_DrImpute", ncomponents = 2) + ggtitle("scone_DrImpute")+ theme(legend.position="none")
p_imp3 <- plotPCASCE(PCA_imp3, colour_by = "cell_line", exprs_values = "SCnorm_DrImpute", ncomponents = 2) + ggtitle("SCnorm_DrImpute")+ theme(legend.position="none")
# p_imp4 <- plotPCASCE(PCA_imp4, colour_by = "cell_line", exprs_values = "BASiCS_DrImpute", ncomponents = 2) + ggtitle("BASiCS_DrImpute")+ theme(legend.position="none")
p_imp5 <- plotPCASCE(PCA_imp5, colour_by = "cell_line", exprs_values = "DESeq_DrImpute", ncomponents = 2) + ggtitle("DESeq_DrImpute")+ theme(legend.position="none")
p_imp6 <- plotPCASCE(PCA_imp6, colour_by = "cell_line", exprs_values = "CPM_DrImpute", ncomponents = 2) + ggtitle("CPM_DrImpute")+ theme(legend.position="none")
p_imp7 <- plotPCASCE(PCA_imp7, colour_by = "cell_line", exprs_values = "TMM_DrImpute", ncomponents = 2) + ggtitle("TMM_DrImpute")+ theme(legend.position="none")
multiplot(p11, p_imp1, p_imp2, p_imp3, p_imp5, p_imp6, p_imp7, cols=3)
```

```{r sil_impu}
sil_imp_1 <- calcSilhouette(PCA_imp1, "linnorm+DrImpute")
sil_imp_2 <- calcSilhouette(PCA_imp2, "scone+DrImpute")
sil_imp_3 <- calcSilhouette(PCA_imp3, "SCnorm+DrImpute")
# sil_imp_4 <- calcSilhouette(PCA_imp4, "BASiCS+DrImpute")
sil_imp_5 <- calcSilhouette(PCA_imp5, "DESeq+DrImpute")
sil_imp_6 <- calcSilhouette(PCA_imp6, "CPM+DrImpute")
sil_imp_7 <- calcSilhouette(PCA_imp7, "TMM+DrImpute")

sil_imp <- rbind(sil_imp_1, sil_imp_2, sil_imp_3, sil_imp_5, sil_imp_6, sil_imp_7)
sil_all_imp <-rbind(sil_all, sil_imp)
```

```{r boxplot_impu}
# pdf("benchmark/silhouette_CEL-seq.pdf")
ggplot(sil_imp, aes(x = reorder(method, sil_width), y=sil_width, fill=reorder(method, sil_width))) + geom_boxplot() + stat_summary(fun.y=mean, geom="point",size=2) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(legend.position = "none")+ labs(x="method", y="silhouette width")
ggplot(sil_all_imp, aes(x = reorder(method, sil_width), y=sil_width, fill=reorder(method, sil_width))) + geom_boxplot() + stat_summary(fun.y=mean, geom="point",size=2) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(legend.position = "none")+ labs(x="method", y="silhouette width")
# dev.off()
```



# F

Function for ANOVA and return F-statistic
```{r}
calculateF <- function(SCE, n=1){
  data <- cbind(reducedDim(SCE), colData(SCE))
  g <- lm(data[,n]~data[,"cell_line"])
  return(anova(g)$F[1])
}
```

```{r}
f <- sapply(c(PCA1, PCA2, PCA3, PCA4, PCA5, PCA6, PCA8, PCA9, PCA10, PCA11, PCA12, PCA13), function(x){calculateF(x)}, simplify = TRUE)
f <- data.frame(f = f, method = c("raw count", "CPM", "scran", "SCnorm", "TMM", "DESeq", "kNN",  "scone", "scran+DrImpute", "scran+SAVER", "Linnorm", "zinbwave"))
```

```{r}
ggplot(f, aes(x = reorder(method,f), y = f, fill=reorder(method,f))) + geom_bar(stat="identity")+ theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(legend.position = "none") + labs(x="method", y="F-statistic")
```

```{r}
f2 <- sapply(c(PCA1, PCA2, PCA3, PCA4, PCA5, PCA6, PCA8, PCA9, PCA10, PCA11, PCA12, PCA13), function(x){calculateF(x, 2)}, simplify = TRUE)
f2 <- data.frame(f = f2, method = c("raw count", "CPM", "scran", "SCnorm", "TMM", "DESeq", "kNN",  "scone", "scran+DrImpute", "scran+SAVER", "Linnorm", "zinbwave"))
```

```{r}
ggplot(f2, aes(x = reorder(method,f), y = f, fill=reorder(method,f))) + geom_bar(stat="identity")+ theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme(legend.position = "none") + labs(x="method", y="F-statistic")
```

```{r}
write.csv(f, "F/sc_10x_1.csv")
write.csv(f2, "F/sc_10x_2.csv")
```

# save

```{r}
write.csv(sil_all, "benchmark/10x_sil_all.csv")
save.image("benchmark_10x.RData")
```